Reproducible Research Course Project 1
======================================
Activity monitoring analysis of an individual over a two month period using fitness tracking device

Data includes the number of steps taken by the individual at 5 minute intervals

First, the file is unzipped and the dataset is read in
     
```{r}
        unzip("activity.zip")
        rawdata <- read.csv("activity.csv", header = TRUE)
        summary(rawdata)
```

Next, we calculate the daily total steps taken and present a summary of the data showing mean and median steps taken

```{r}
        daily_sums <- aggregate(steps ~ date, rawdata, sum, na.rm=TRUE)
        summary(daily_sums$steps)
```

Below is a histogram of the daily steps taken showing the frequency distribution

```{r}
        hist(daily_sums$steps,
                main = "Daily Steps Frequency",
                col = "blue",
                xlab = "Daily Steps Taken",
                ylab = "Frequency"
     )
```

To examine the average daily activity patterns, the below aggregates the data across 5 minute intervals using the mean and presents the data in a time series plot

```{r}
        interval_means <- aggregate(steps ~ interval, rawdata, mean, na.rm = TRUE)
        plot(interval_means$interval, interval_means$steps,
             type = "l",
             main = "Average Steps Taken by 5-minute Interval",
             xlab = "Interval",
             ylab = "Average Steps Taken",
             col = "green"
             )
        interval_means[which.max(interval_means$steps),]
```

To deal with missing values, the below code creates a dataframe of incomplete values and prints the number of missing values

```{r}
  NA_values <- rawdata[(is.na(rawdata)),]
  nrow(NA_values)
```

Then we make a table of the average steps taken by interval for all complete observations, and merge this table with the incomplete observations, creating a new column for steps that replaces the NA values with that interval's average

```{r}
average_interval <- aggregate(steps ~ interval, rawdata, mean, na.rm=TRUE)
colnames(average_interval) = c("interval", "steps.computed")
replaced_NA <- merge(NA_values, average_interval, by="interval", all.x = TRUE)
```

Below, a dataset of just complete values is created, along with a new column replicating the steps column, which will contain all modified data once we merge the datasets

```{r}
nomissing <- rawdata[complete.cases(rawdata),]
nomissing$steps.computed <- nomissing$steps
activity_data <- rbind(replaced_NA, nomissing)
activity_data$steps.computed <- round(activity_data$steps.computed, digits = 0)
```

With the modified dataset with no missing values, we recreate the above histogram

```{r}
daily_sums2 <- aggregate(steps.computed ~ date, activity_data, sum)
hist(daily_sums2$steps.computed,
                main = "Daily Steps Frequency",
                col = "red",
                xlab = "Daily Steps Taken",
                ylab = "Frequency"
 )
        summary(daily_sums2$steps.computed)
```

